---
title: "Pittsburgh Pirates Application"
author: "Liam Jennings"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

Oneil Cruz is a young and upcoming player in the Pirates organization.
I saw a few Pirates games in person, and watched about half of them from my TV at home in Maryland.
Oneil Cruz can hit the ball harder than almost anyone. He hit the hardest hit ball in the Statcast Era (since 2015)
at 122.4 mph. However, he finished with a line of .233/.294/.450 
with 126 strikeouts (34.9% strikeout percentage) and 28 walks (7.8% walk percentage). 
The league average rates were 22.4% and 8.4%, respectively.
Oneil Cruz is below average in both categories. It's one of his weaknesses (albeit, he's young
and has plenty of time and room to grow). There's one series that stood out in particular.
The Pirates played against the Orioles in Baltimore on August 5th - August 7th. My dad and I went to the first game of the series. Oneil Cruz struck out three times. The first one was on three pitches. The second one was on a slider in the dirt. The third strikeout was against Orioles' Félix Bautista, a dominant reliever with a nasty splitter. The Pirates lost 1 - 0. For the rest of the series, Oneil Cruz struck out seven times, three against lefties. All three pitches against the southpaws were sliders.
 
Question: How does Oneil fair against lefties? How about off-speed pitches from lefties?
Is it just sliders?
Also, how do you set up Cruz for a strikeout? As a pitcher, what is your pitch sequence against Cruz?

The following information will be useful for working with the pitch-by-pitch data

Pitch Type Abbreviations:

FASTBALLS:

- FF = four-seam fastball (most common pitch in baseball)

- FT = two-seam fastball (more movement than FF)

- FC = cutter

- SI = sinker

OFFSPEED:

- SL = slider 

- CH = changeup

- CU = curveball

- KC = knuckle-curve

- FS - Splitter / split-fingered

- KN = knuckleball

- EP = eephus

Pitch locations:

- plate_x = the left/right distance, in feet, of the pitch 
          from the middle of the plate as it crossed home plate from
          the catcher's point of view with distances to the right as 
          positive, left as negative.
          
- plate_z = height of the pitch in feet as it crossed the front of home plate

# Loading in the data and libraries
Load in libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(ggrepel)
```

Load in the Oneil Cruz (id: 458015) data:
```{r}
cruz_data <- read.csv("oneil_cruz_2022.csv")
```

Load in all 2022 Statcast data:
```{r}
statcast_2022 <- read.csv("statcast_2022.csv")
```

Create a variable for the count and strikeout counter
```{r}
cruz_data <- cruz_data %>%
  mutate(count = paste(as.character(balls), "-", as.character(strikes))) %>%
  mutate(strikeout = ifelse(events == "strikeout" | events == "strikeout_double_play", 1, 0))

statcast_2022 <- statcast_2022 %>%
  mutate(strikeout = ifelse(events == "strikeout" | events == "strikeout_double_play", 1, 0))
```

Create a dataset with every at-bat where Oneil Cruz struck out
```{r}
cruz_strikeouts <- cruz_data %>%
  filter(des %in% c("Oneil Cruz called out on strikes.", "Oneil Cruz strikes out swinging.", 
                    "Oneil Cruz strikes out swinging and Jack Suwinski caught stealing 2nd, catcher Michael Papierski to third baseman Brandon Drury to catcher Michael Papierski.",
                    "Oneil Cruz strikes out on a foul tip.", "Oneil Cruz strikes out swinging.  Rodolfo Castro steals (3) 2nd base.")) %>%
  select(game_date, inning, pitch_type, description, zone,
         count, plate_x, plate_z, p_throws, pitch_number, sz_top, sz_bot, events, strikeout) %>%
  arrange(game_date, inning)
```

# Location of Pitches

We're going to look at pitches thrown against Cruz on his average strike zone.


## Oneil Cruz's average strike zone
```{r}
# Outside layer
top_zone <- mean(cruz_data$sz_top)
bot_zone <- mean(cruz_data$sz_bot)
left_zone <- -0.95
right_zone <- 0.95
strike_zone_df <- data.frame(
  x = c(left_zone, left_zone, right_zone, right_zone, left_zone),
  y = c(bot_zone, top_zone, top_zone, bot_zone, bot_zone)
)

# Horizontal lines to break the strike zone into ninths
top_zone <- mean(cruz_data$sz_top)
bot_zone <- mean(cruz_data$sz_bot)
left_third <- -0.95/3
right_third <- 0.95/3
inside_strike_zone_df <- data.frame(
  x = c(left_third, left_third, right_third, right_third, left_third),
  y = c(bot_zone, top_zone, top_zone, bot_zone, bot_zone)
)

# Vertical lines to break the strike zone into ninths
top_third <- top_zone - ((top_zone - bot_zone)/3)
bot_third <- ((top_zone - bot_zone)/3) + bot_zone
vertical_strike_zone_df <- data.frame(
  x = c(left_zone, left_zone, right_zone, right_zone, left_zone),
  y = c(bot_third, top_third, top_third, bot_third, bot_third)
)
```

## Location of pitches on strikeouts
### August 5th, 2022
#### 2nd Inning
```{r, echo=FALSE}
cruz_strikeouts %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FS", "FT", "KC", "SI", "SL"), 
         game_date == "2022-08-05", inning == 2) %>%
  ggplot(aes(x = plate_x, y = plate_z, color = pitch_type)) + 
  geom_point(alpha = 0.75) + 
  geom_path(data = strike_zone_df, aes(x, y), linewidth = 1.5, color = "black") + 
  geom_path(data = inside_strike_zone_df, aes(x, y), color = "black") +
  geom_path(data = vertical_strike_zone_df, aes(x, y), color = "black") +
  geom_text_repel(aes(label = count)) +
  coord_fixed() +
  labs(title = "Location of Pitches Thrown Against Oneil Cruz on August 5th by Pitch Type",
       caption = "Data courtesy of MLBAM",
       x = "Horizontal Location (feet)",
       y = "Vertical Location (feet)", 
       color = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.50))
```

The graph shows the pitches Oneil Cruz saw (from the catcher's perspective) on August 5th, 2022,
in the 2nd inning. The graph is labeled by pitch type (color) and count (label on the point). 
The first pitch was a changeup just to the left of the middle of the plate.
The second pitch was a 4-seam fastball nearly in the heart of the plate. The final pitch of the
at-bat was curveball almost in the dirt, which was a swinging strike and strikeout for Cruz.
Oneil Cruz's strike zone can be shown for more than just one at-bat.

### August 5th, 2022
#### Entire Game
```{r, echo=FALSE}
cruz_strikeouts %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FS", "FT", "KC", "SI", "SL"), 
         game_date == "2022-08-05") %>%
  ggplot(aes(x = plate_x, y = plate_z, color = pitch_type)) + 
  geom_point(alpha = 0.75) + 
  geom_path(data = strike_zone_df, aes(x, y), linewidth = 1.5, color = "black") + 
  geom_path(data = inside_strike_zone_df, aes(x, y), color = "black") +
  geom_path(data = vertical_strike_zone_df, aes(x, y), color = "black") +
  geom_text_repel(aes(label = count)) +
  coord_fixed() +
  labs(title = "Location of Pitches Thrown Against Oneil Cruz on August 5th by Pitch Type",
       caption = "Data courtesy of MLBAM",
       x = "Horizontal Location (feet)",
       y = "Vertical Location (feet)", 
       color = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

Of all of the at-bats where Oneil Cruz struck out, he saw only four pitches in the strike zone,
which was 25% of the pitches. He struck out three times. The three 
pitches he struck out on were a curveball, a slider, and a splitter. The curveball
was roughly a foot below his average strike zone. The 1 - 2 slider was nearly in the 
dirt (less than six inches off the ground) and over a foot and a half off the plate. 
The other two sliders that were thrown against him this game were the second or third
pitches of the at-bat (neither one on two strike counts) and below the zone.
Is that a common location for sliders against Oneil Cruz?

### Series against Baltimore (August 5th - 7th)
```{r, echo=FALSE}
cruz_strikeouts %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FS", "FT", "KC", "SI", "SL"), 
         game_date == "2022-08-05" | game_date ==  "2022-08-06" | game_date ==  "2022-08-07") %>%
  ggplot(aes(x = plate_x, y = plate_z, color = pitch_type)) + 
  geom_point(alpha = 0.75) + 
  geom_path(data = strike_zone_df, aes(x, y), linewidth = 1.5, color = "black") + 
  geom_path(data = inside_strike_zone_df, aes(x, y), color = "black") +
  geom_path(data = vertical_strike_zone_df, aes(x, y), color = "black") +
  geom_text_repel(aes(label = count)) +
  coord_fixed() +
  labs(title = "Location of Pitches Thrown Against Oneil Cruz on Aug 5th - 7th by Pitch Type",
       caption = "Data courtesy of MLBAM",
       x = "Horizontal Location (feet)",
       y = "Vertical Location (feet)", 
       color = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

On at-bats when Oneil Cruz struck out, pitchers threw him eight sliders (~27% of pitches).
Seven out of eight of them were below the zone. Three of those eight sliders 
were on two strike counts, which he subsequently struck out on.
Three of them were on 0 - 1 counts, the second pitch of the at-bat. The other 
two were on 1 - 1 counts. In this series, pitchers were throwing Cruz sliders to the opposite
side of the plate as him and below the zone. Additionally, five out of the eight
were not on two strike counts. 

### Sliders in at-bats ending in strikeouts over the whole season 
```{r, echo=FALSE}
cruz_strikeouts %>%
  filter(pitch_type == "SL") %>%
  ggplot(aes(x = plate_x, y = plate_z)) + 
  geom_point(aes(color = as.factor(zone)), alpha = 0.75) + 
  geom_path(data = strike_zone_df, aes(x, y), linewidth = 1.5, color = "black") + 
  geom_path(data = inside_strike_zone_df, aes(x, y), color = "black") +
  geom_path(data = vertical_strike_zone_df, aes(x, y), color = "black") +
  coord_fixed() +
  labs(title = "Sliders Thrown Against Oneil Cruz in 2022",
       caption = "Data courtesy of MLBAM",
       x = "Horizontal Location (feet)",
       y = "Vertical Location (feet)",
       color = "Zones") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

There appears to be a trend of sliders being thrown below, to the left (catcher's perspective), 
and in the bottom third of the zone when considering all sliders thrown against 
Oneil Cruz in ABs ending in a strikeout. Those locations are hardest for Oneil 
Cruz to hit as a left-handed hitter. After looking at a couple of at-bats
and general location of sliders against Oneil Cruz, we begin to look deeper. 

# What pitches did Cruz face this season?

We're going to look at the number of different types of pitches Cruz faced this season.
Then, we're going to compare what he was thrown against other left-handed hitters.

## Types of pitches thrown against Oneil Cruz in 2022
```{r, echo=FALSE}
ggplot(cruz_data, 
       aes(x = pitch_type,
           y = after_stat(count)/sum(after_stat(count)))) +
  geom_bar(aes(fill = pitch_type)) +
  scale_fill_brewer(palette = "Set1", guide = 'none') +
  labs(title = "Types of Pitches Thrown Against Oneil Cruz in 2022",
       x = "Pitch Type",
       y = "Proportion of Pitches",
       caption = "Data courtesy of MLBAM") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


24.2% of pitches Cruz saw in 2022 were 4-seam fastballs. The second most pitches
he saw were sliders at 21.7%. How does that compare to other left-handed hitters?

## Compared to ALL left-handed hitters
```{r, echo=FALSE}
statcast_2022 %>%
  filter(player_name != "Cruz, Oneil", stand == "L", pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "FS", "KC", "SI", "SL")) %>%
  ggplot(aes(x = pitch_type,
           y = after_stat(count)/sum(after_stat(count)))) +
  geom_bar(aes(fill = pitch_type)) +
  scale_fill_brewer(palette = "Set1", guide = 'none') +
  labs(title = "Types of Pitches Thrown Against Left-Handed Hitters in 2022",
       x = "Pitch Type",
       y = "Proportion of Pitches",
       caption = "Data courtesy of MLBAM") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Other left-handed batters are seeing sliders at the second highest clip as well; however,
the number is approximately 16%, almost 6% lower than Oneil Cruz's rate of sliders.
Additionally, Cruz sees more than 10% less 4-seam fastballs than other lefties. 
He is seeing an above average amount of sliders compared to other left-handed batters.
Pitchers are throwing a lot of sliders Cruz's way. 

# Strikeout Data

Next, we're looking at the number of pitches Cruz strikes out against and comparing
that to other left-handed batters.

## Number of strikeouts by each pitch
```{r, echo=FALSE}
cruz_strikeouts %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "KC", "SI", "FS", "SL")) %>%
  group_by(pitch_type) %>%
  summarize(n_strikeouts = sum(strikeout, na.rm = TRUE)) %>%
  mutate(SO_rate = n_strikeouts / sum(n_strikeouts)) %>%
  ggplot(aes(x = pitch_type, y = SO_rate, fill = pitch_type)) + 
  scale_fill_brewer(palette = "Set2", guide = 'none') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_bar(stat = "identity") + 
  labs(title = "Oneil Cruz Strikeout by Pitch Type in 2022",
       caption = "Data courtesy of MLBAM",
       x = "Pitch Type",
       y = "Strikeout %") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Oneil Cruz struck out 126 times last year. The pitch that he struck out the most 
on was slider (44 times or 35%), which is twice as much as the second most struck out
pitch, the changeup (20 times or ~16%).

## Strikeout pitches compared to other left-handed batters
```{r, echo=FALSE}
statcast_2022 %>%
  filter(player_name != "Cruz, Oneil" ,pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "KC", "SI", "FS", "SL") & stand == "L") %>%
  group_by(pitch_type) %>%
  summarize(n_strikeouts = sum(strikeout, na.rm = TRUE)) %>%
  mutate(SO_rate = n_strikeouts / sum(n_strikeouts)) %>%
  ggplot(aes(x = pitch_type, y = SO_rate , fill = pitch_type)) + 
  scale_fill_brewer(palette = "Set2", guide = 'none') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_bar(stat = "identity") +
  labs(title = "Left Handed Hitters Strikeout by Pitch Type in 2022",
       caption = "Data courtesy of MLBAM",
       x = "Pitch Type",
       y = "Strikeout %") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Other left handed batters' biggest strikeout pitch is the 4-seam fastball, 
the cause of over 30% of strikeouts. Oneil Cruz struck out on 4-seam fastballs
less than 15% of the time, which is less than half of what other lefties were doing. 
The slider was the second biggest strikeout pitch against left-handed hitters at
a rate of approximately 24%, which is over 11% lower than Cruz's rate against sliders.


# How do pitchers set Oneil Cruz up for sliders?

Oneil Cruz strikes out against sliders the most. We're going to look at the pitch 
sequencing on at-bats where he strikes out on a slider. 

```{r, echo=FALSE}
cruz_sliders <- cruz_strikeouts[c(1:16, 21:23, 41:53, 74:76, 101:106, 122:143, 147:149, 165:168, 
                                  177:181, 230:233, 239:245, 266:269, 274:301, 306:309, 316:323,
                                  330:333, 350:356, 382:385, 391:399, 415:426, 442:459, 466:472,
                                  543:550, 556:559), ]
```


## Pitch Sequencing on Slider Strikeouts
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL")) %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = pitches_thrown, fill = pitch_type)) + 
  geom_bar(position = "dodge", stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "% of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

##  Frequency of Pitch Sequencing on Slider Strikeouts
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "L") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = n_pitches, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO Frequency",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "# of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


On the first pitch of an at-bat where Cruz struck out an a slider, pitchers threw a slider 45% of the time. 
Pitchers are throwing sliders the most, regardless of the count. On 0 - 1 and 1 - 0, the second
pitch of an at-bat, pitchers are throwing a slider over 50% in each count, the most of any pitch.
Whether pitchers are ahead, even, or behind in the count (except 2-0 and 2-1), they are 
throwing sliders the most. Even in 3 - 1 counts, a typical fastball count, pitchers are 
throwing Cruz a slider (2/2 on 3 - 1 counts). On at-bats where Oneil Cruz struck out 
on sliders, pitchers tend to be ahead of him in the count. There was only one instance
where a pitcher was down 2-0, six instances of 2-1, two instances of 3-1, and twelve instances of 
1-0. 

# Left vs. Right handed pitchers

Now, we're going to see if there's a difference between Cruz vs. 
right- and left-handed pitching. We're going to compare his general pitch sequencing
to his pitch sequencing on at-bats ending in a slider strike out. 

## LHPs

First, we're looking at LHPs, the side that I believe he is worse against.

### Percentage of pitch type by count on at-bats ending in slider strikeouts
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "L") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = pitches_thrown, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO vs. LHP",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "% of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


### Frequency of pitch type by count on at-bats ending in slider strikeouts
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "L") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = n_pitches, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO vs. LHP Frequency",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "# of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Lefties are throwing sliders the most in every count except 2 - 0 and 2 - 1 
where the end result was an Oneil Cruz strikeout on a slider. Over 50% of the first
pitches were a slider. The second most thrown pitch was a 4-seam fastball at a 
quarter of first pitches. In 0 - 1 counts, left-handed pitchers are throwing 
sliders over 75% of the time. In 1 - 0 counts, sliders are thrown in
approximately 60% of at-bats. In 1 - 1 counts, slider is the pitch 50% of the time
Typically, off-speed pitches are used later in the count; however, left-handed
pitchers are setting him up with sliders. Additionally, LHPs are throwing sinkers
the second most in six counts. Sinkers break towards a batter like a 2-seam fastball;
however, there is significant downward drop or "sinking" movement. Sliders
have the opposite lateral movement of a sinker; however, they drop, but not as much
(depends on the grip, arm action, pitcher, etc.). For most batters, it is easier 
to bat against a pitcher of the opposite handedness because it is easier to see pitches
break. Oneil Cruz may struggle to see pitches breaking when batting against 
lefties. 

### Compare to general pitch sequencing against LHPs
```{r, echo=FALSE}
cruz_data %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "L") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = pitches_thrown, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Oneil Cruz Pitch Sequencing vs. LHP by % of Pitch Type",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "% of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


### Frequency of pitches against LHPs
```{r, echo=FALSE}
cruz_data %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "L") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = n_pitches, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing vs. LHP Number of Pitches",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "# of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Against lefties, Oneil Cruz sees a slider approximately 40% of the time on his
first pitch, the most by almost twice as much. On 0 - 1 counts, he sees a 
slider on over half of the pitches, and no other pitch is above 20%. 1 - 0 and 1 - 1
are very similar to 0 - 0, except on 1 - 0, roughly 35% of the pitches are sliders
and on 1 - 1, a little less than 45% of pitches are sliders. All four of those counts
are almost identical to the percentages in at-bats where he strikes out on a slider; therefore,
that sample represents the population well. There was only one instance of a 
2 - 0 count in slider strikeout ABs, but there were 13 total times against LHPs.
Sinkers were thrown 4 out of those 13 times (~31%), but sliders were tied second
with 3 out of 13 times (~21%). On 2 - 2 counts, sliders were thrown about 29% of 
the time, which is similar to the percentage in slider strikeout at-bats; however, 
it's thrown the most against LHPs in general. There were only three times 
when Oneil Cruz reached a 3 - 0 count against lefties, and a slider was not thrown once. 
Of the nine 3 - 1 counts, sliders were thrown a third of the time, tied for first
with curveball. It's surprising that pitchers are trying to get Oneil Cruz to 
swing and miss on a 3 - 1 count instead of a fastball. Sliders are the most thrown or 
tied for most thrown (3 - 2) on all two strike counts. The highest percentages
of sliders thrown during a count is slightly over 50% of the time on 0 - 1 counts.

## RHPs

Next, we take a look at his pitch sequencing against RHPs. 

### Percentage of pitch type by count on at-bats ending in slider SOs 
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL") & p_throws == "R") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = pitches_thrown, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO vs. RHP",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "% of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


### Frequency of pitch type by count on at-bats ending in slider SOs
```{r, echo=FALSE}
cruz_sliders %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "R") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = n_pitches, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing on At-Bats with Slider SO vs. RHP Frequency",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "# of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Right-handed pitchers are not throwing the same number of sliders earlier 
in the count as often as lefties are.
Sliders are the primary pitch in six counts and are thrown over 50% of the time in
all of them compared to seven counts and at least 50% of the time for all of them for lefties.
The biggest differences are in the 0 - 0, 0 - 1, 1 - 0, and 2 - 1 counts. 
RHPs only throw sliders 25% of the time on the first pitch in slider strikeout at-bats.
Also, instead of lefties throwing a slider three quarters of the time on 0 - 1,
righties throw it just over 50%, which is still a high number. On 1 - 0 counts, 
right-handed pitchers exclusively throw 4-seam fastballs. However, when it is 
2 - 1, sliders are thrown 50% of the time, tied with 4-seam fastballs. 
When right-handed pitchers go down in the count after the first pitch,
they tried to even up the count with a 4-seam fastball every time instead of an off-speed pitch.
On 2 - 1, pitchers weren't afraid to throw an off-speed pitch to either get a chase out
of the zone or a whiff. This strategy proves successful because RHPs never had a
3 - 1 count against Oneil Cruz when he struck out on a slider. 

### Compare to general pitch sequencing against RHPs
```{r, echo=FALSE}
cruz_data %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "R") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = pitches_thrown, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Oneil Cruz Pitch Sequencing vs. RHP by % of Pitch Type",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "% of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


### Frequency of pitches against RHPs
```{r, echo=FALSE}
cruz_data %>%
  filter(pitch_type %in% c("CH", "CU", "FC", "FF", "FT", "SI", "SL"), p_throws == "R") %>%
  group_by(count, pitch_type) %>%
  summarize(n_pitches = n()) %>%
  mutate(pitches_thrown = n_pitches / sum(n_pitches)) %>%
  ggplot(aes(x = count, y = n_pitches, fill = pitch_type)) + 
  geom_bar(position = position_dodge(.75), stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Oneil Cruz Pitch Sequencing vs. RHP Number of Pitches",
       caption = "Data courtesy of MLBAM",
       x = "Count",
       y = "# of Pitches Thrown", 
       fill = "Pitch Type") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


Against righties, Oneil Cruz sees a 4-seam fastball approximately 30% of the time on his
first pitch, the most by over 10%. His first pitch slider rate is under 10%. 
On 0 - 1 counts, he sees a 4-seam fastball a little less than 30% of the time, which
is over 20% below the rate against RHPs when he strikes out on a slider. 
On 1 - 0, 2 - 0, and 2 - 1, he sees 4-seam fastballs at a rate around 30%, which is the most
of all the pitches. On 1 - 0, sliders are thrown at the lowest rate (less than 10%).
For 2 - 0, sliders are thrown at the same rate as curveballs, which ties them
for first among offspeed pitches. Cruz sees sliders at a rate of approximately 15% 
on 2 - 1 counts, which is good for fourth among pitch types. Cruz was not thrown 
a single slider in 3 - 0 counts. Changeups are thrown the most, a little over a 
quarter, on 1 - 1 counts, and 4-seam fastballs are second. 3 - 1 counts are similar; however, 
only changeups are tossed over 20% of the time, and slider rate is ranked last with
curveballs (~11%). On two strike counts, changeups are favored when the count is 
1 - 2 or 3 - 2, 4-seam fastballs are thrown most on 0 - 2, and sliders are tied with
4-seam fastballs on 2 - 2 counts. 2 - 2 counts is the only count where sliders 
hold a share of being thrown the most. There are only two counts where sliders are
seen more than 20% of the time, 0 - 2 and 2 - 2. The slider strikeouts against righties
do not indicate a trend of pitch sequencing vs. right-handed pitching. 


# Conclusion
The first thing we looked at was the proportion of pitches Oneil Cruz faced this season. He saw a slider on 21.7% of pitches, second to a 4-seam fastball 24.2% of the time. It appears that he saw an abnormal amount of sliders. I tested that claim by looking at the proportion of pitches thrown against all left-handed hitters. Cruz sees almost 6% more sliders and 10% less 4-seam fastballs than the average lefty. The pitches he saw last season were different than the average left-handed hitter; therefore, the next step was to investigate the number of strikeouts by each pitch and compare it to the rest of the population. Oneil Cruz strikes out the most against sliders (35% of strikeout pitches). Not a single other pitch was above 17%. Left-handed hitters struck out on 4-seam fastballs over 30% of the time, which was the highest strike out pitch for lefties. Oneil Cruz struck out on 4-seam fastballs less than 15% of the time, which is more than half of other lefties. The slider was the second biggest strikeout pitch against left-handed hitters at a rate of approximately 24%, which is over 11% lower than Cruz’s rate against sliders. After determining that Oneil Cruz strikes out the most against sliders, I decided to look at how pitchers set him up for sliders. I broke down each pitch type into the percentage thrown in a count on at-bats that end in a slider strikeout. Pitchers are throwing sliders almost half of the time on the first pitch. Additionally, Cruz sees sliders over 50% of the time on 0 - 1, 1 - 0, and 1 - 1 counts. Sliders are being thrown early in the count to set up sliders later in the count. However, my next question was if there was a difference between Cruz's pitch sequencing against LHPs vs. RHPs. The conclusion is that lefties are throwing more sliders than righties against Oneil Cruz. On at-bats that end in a slider strikeout, LHPs are throwing sliders at or greater than half of the time in 0 - 0, 0 - 1, 1 - 0, and 1 - 1 counts. Compared to his general pitch sequencing against lefties, sliders are thrown the most in every count except 2 - 0 and 3 - 0. The same trend cannot be said for Cruz vs. RHPs. On at-bats where Cruz struck out on a slider against RHPs, sliders were thrown over 50% of the time on 0 - 1 and 1 - 1 counts. Although, this trend doesn't carry over to general pitch sequencing against Cruz. Sliders were only thrown the most in one count (2 - 2, tied with 4-seam fastball). RHPs were not throwing sliders very often against Cruz, especially early in the count. The slider rate in 0 - 0 and 1 - 0 counts were below 10% and only two counts had a rate higher than 20%, 0 - 2 and 2 - 2, both two strike counts. RHPs were not using sliders to set up Oneil Cruz. In every count, lefties are throwing more sliders than righties, proportionally. LHPs appear to be effective against Oneil Cruz when using sliders, regardless of the count. It is the pitch thrown the most early in the count (most in 0 - 0, 0 - 1, 1 - 0, and 1 - 1) and in two strike counts. Cruz strikes out more against left-handed pitchers with a 53% strikeout rate compared to a 26.5% strikeout rate against right-handed pitchers. Lefties are setting up Oneil Cruz with sliders and will likely continue this strategy as long as success is maintained. 
